---
title: "SDL_R_NMF"
author: "Delin Sun"
date: "6/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Statistical analyses of the NMF outputs (7/9/15-components solutions by Viraj) of cortical thickness data from NCANDA binge drinking project

Delin Sun
Version date 07/30/2020.

### Independent Variables

Drinking_class: Severity of drinking problem. 0 = no/to low drinkers, 1 = medium drinkers, 2= heavy drinkers, 3 = heavy drinker + binge. We can also combine 2 and 3 as heavy drinking.

Age_d: The difference between a subject’s age at each scan and their mean age across individual timepoints (i.e. scan_age - subject_mean_age).

Age_m: the difference between a subject’s mean age across visits and the mean age of the entire sample across timepoints (i.e. subject_mean_age - mean(subject_mean_age)), thus centering cohort age at the sample mean. Each participant’s cohort age remained constant across timepoints. 

life_trauma_RP: Cumulative trauma at baseline. It was quantified as the sum of reported DSM-IV or 5 Criterion A traumatic events of trauma history. DSM Criterion A traumas and PTSD symptoms were collected as part of the baseline assessment from the youth, and parent with the Computerized Semi-Structured Assessment for the Genetics of Alcoholism (SSAGA) (Bucholz, Cadoret et al. 1994, Hesselbrock, Easton et al. 1999). The SSAGA was modified for adolescents and includes an interview assessment of DSM-IV and 5 Axis I disorders, DSM Criterion A traumas and PTSD interview questions (Brown, Brumback et al. 2015). A traumatic event was counted once if either the parent and/or youth reported an event. No participant met DSM criteria for PTSD. Its range is 0~5. We can combine 4 and 5 as a group.

Fh_alc_density: Family history of alcohol use and dependence (AUD) density. It was calculated based on the presence of AUD in first and second-degree relatives (Rice, Reich et al. 1995). Its range is 0~4.

Ses: Socio economic status quantified using the highest parental years of education of either parent (Brown, Myers et al. 1998). The range is 6~20 yrs. We can categorize 0 = 6~12 yrs (high school or less), 1 = 13~15 yrs (associate), 2 = 16 yrs (BA, BS), 3 = 16 plus yrs (masters or higher). Or we can do 0 = 6~12 yrs, 1 = 13~20 yrs.

Race: Self-identified race/ethnicity instead of “race”. Totally 3 categories: African-American/Black (AAmer), Caucasian/White (Cauc), and Other.

Sex: Male and Female.

Subject: Participant ID as a random effect variable.


## Set path and load packages
```{r,echo=FALSE,warning=FALSE,include=FALSE}
#setwd("C:/Users/sunde/Documents/Projects/Viraj/baseline_nmf_glm_analysis")
library("readxl") # load .xls or .xlsx files
library(tidyverse) # for data manipulation
library(psych)
library(lme4) # lm/lmer models
library(lmerTest)# p-value that are not included in lme4
library(ggplot2)
library(sjPlot) # plot interactions of lm/lmer outputs
library(ggpubr) # plot multiple panels
library(plot.matrix)
```

## Define functions
Totally 4 functions
--- SDL_lm, extract models/coefficients/p values/plots for each component
Already cleaned by Viraj's Matlab scripts
--- SDL_heat, heatmap of p values across factors and componnets
--- SDL_plt,  plot multiple main/interaction effects
--- SDL_all,  call the above 3 functions
```{r,echo=FALSE}

#######################################################################
# a function to model (lm/lmer/glm) data for each ROI, and extract models/coefficients/p-values of models
# Input
# --- y, dataframe column, e.g. df$col01, or select(col1:col10)
# --- df, dataframe
# --- Otype, the type of outputs, e.g. "model", "coef", and "p"
# --- ftxt, text for the formula, e.g. "lmer(formula = y ~ age_d + age_m + drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject)"
# --- fterms, which factor to be plotted
# --- fxy, x and y labels
# Output
# --- model, list of models
# --- coef,  matrix of coefficients
# --- pval,  matrix of coefficients
# --- plt,   list of plotting objects
SDL_lm <- function(y,df,Otype,ftxt,fterms=NULL,fxy=NULL){
  model <- eval(parse(text=paste(substr(ftxt,1,nchar(ftxt)-1),", data = df)")))
  coef  <- model %>% summary() %>% coefficients()
  pval  <- coef  %>% .[,"Pr(>|t|)"]
  plt   <- model %>% plot_model(type = "pred", terms = fterms, title = "", axis.title = fxy)
  switch(Otype,"model"=return(model),"coef"=return(list(coef)),"pval"=return(pval),"plt"=return(list(plt)))
}
##############################################################################

##############################################################################
# A function to plot the heatmap of statistical significance across factors and components
# Input
# --- Pval, p-value matrix
# Output
# --- a heatmap of statistical significance
SDL_heat <- function(Pval){
  heatmap(Pval, margins =c(7,10), Colv=NA, Rowv=NA, main="p-values", scale="none", breaks=c(0,0.001,0.01,0.05,1), col=c('yellow','orange','red','grey')) # no color bar
}
##############################################################################

##############################################################################
# A function to plot the significant main and interaction effects
# Input
# --- Plt, plot list made by SDL_lm function
# --- Pval, p-value matrix made by SDL_lm function
# --- ffact, the factor used to threshold p-value matrix
# --- pth, the significance threshold
# Output
# --- a heatmap of statistical significance
SDL_plt <- function(Plt,Pval,ffact='drinking_class',pth=0.05){
  if (length(which(Pval[ffact,]<pth))){
  ind <- Pval[ffact,] < pth # index of significant components
  Plt <- Plt[ind] # remove non-significant components
  flabel <- colnames(Pval[,ind]) # labels of the significant components
  ggarrange(plotlist=Plt,labels=flabel,common.legend=T) # multiple plots
  }else{
    paste0("NS: ",ffact)
  }
}
##############################################################################



##############################################################################
# A function to combining the aforementioned 3 functions
# Input
# --- col, column(s) of dependent variables, e.g. df%>%select(c1:c7)
# --- df, dataframe
# --- fcorr, the method for correction of multiple comparisons, e.g. 'fdr','bonferroni','none'
# --- ftxt, lm/lmer/glm formula
# --- fterms, independent variables
# --- fxy, plot's x, y labels
# --- ffact, the factor used to threshold p-value matrix
# Output
# --- a heatmap of statistical significance
SDL_all <- function(col,df,fcorr,ftxt,fterms,fxy,ffact){
  #get p-value matrix
  Pval <- col %>% sapply(SDL_lm, df=df, Otype="pval", ftxt=ftxt)  
  Pval_fdr <- Pval %>% t() %>% as.data.frame() %>% sapply(p.adjust,method=fcorr) %>% t()
  colnames(Pval_fdr) <- colnames(Pval)
  
  # plot significance heatmap
  Pval_fdr %>% SDL_heat() # show all significant effects (p < 0.05 unc.)

  # plot significant main or interaction effects
  col %>% sapply(SDL_lm, df=df, Otype="plt", ftxt=ftxt, fterms=fterms, fxy=fxy) %>% SDL_plt(Pval=Pval_fdr,ffact=ffact, pth=0.05) # statistical threshold
}
##############################################################################

```

# Data cleaning
Only employ subjects whose drinking class = 0 at baseline

```{r,echo=FALSE}
#df0 <- read.csv("Data_cleaned.csv",header=T)

 # raw data (demographic doc), 2815 observations
df0 <- read_excel("../original/ncanda_all_sites_out_drkclss_20200617.xlsx")

# print number of subjects per site
df0 %>% select(c('subject','visit','drinking_class','site')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% group_by(site) %>% summarise(n=n())

 # NMF outputs, 2809 observations
dfx <- read.csv("./11.03/NMF_k07_d20484_n2809.csv",header=T) %>% rename(folder = sbj) # 7-pattern solution
#dfx <- read.csv("./11.03/NMF_k09_d20484_n2809.csv",header=T) %>% rename(folder = sbj) # 9-pattern solution

# merge demo info & NMF outputs, 2809 observations
df0 <- merge(df0,dfx,by="folder",all.y=T)


# print number of subjects per site
df0 %>% select(c('subject','visit','drinking_class','site')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% group_by(site) %>% summarise(n=n())

 # print number of subjects per drinking_class level at baseline
df0 %>% select(c('subject','visit','drinking_class')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% group_by(baseline) %>% summarise(n=n())


# Only keep those drinking_class = 0 at baseline, 657 subjects
df1 <- df0 %>% select(c('subject','visit','drinking_class')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% filter(baseline==0) # keep those with drinking_class=0 at baseline

# 657 subjects correspond to 2628 observations
df1 <- df1 %>% pivot_longer(-subject, names_to = "visit", values_to = "drinking_class")

# final data for analyses, 2628 observations
df <- merge(df0,df1,by=c("subject","visit"),all.y=T) %>% rename(drinking_class=drinking_class.y) 

# total individual study visits with drinking_class values, 2253 observations
df %>% summarise(n=n())

# calculate age_m & age_d
df1 <- df %>% select(c('subject','visit','mri_t1_age')) %>% pivot_wider(names_from = visit, values_from = mri_t1_age) %>% mutate(subject_mean_age = rowMeans(select(., baseline:followup_3y),na.rm=T), age_m = subject_mean_age - mean(subject_mean_age), baseline=baseline-subject_mean_age, followup_1y=followup_1y-subject_mean_age, followup_2y=followup_2y-subject_mean_age, followup_3y=followup_3y-subject_mean_age) # calculate subject_mean_age, age_m and age_d (values are in each of visits)

# add age_d to df
dfx <- df1 %>% select(subject:followup_3y) %>% pivot_longer(-subject, names_to = "visit", values_to = "age_d") # turn to long table
df <- merge(df,dfx,by=c("subject","visit"),all.y=T) 

# add age_m to df
dfx <- df1 %>% mutate(baseline=age_m, followup_1y=age_m, followup_2y=age_m, followup_3y=age_m) %>% select(subject:followup_3y) %>% pivot_longer(-subject, names_to = "visit", values_to = "age_m") # turn to long table
df <- merge(df,dfx,by=c("subject","visit"),all.y=T)

# total individual study visits, 2628 observations
df %>% summarise(n=n())

# total individual study visits with drinking_class values, 2253 observations
df %>% filter(drinking_class>=0) %>% summarise(n=n())

```

# Statistical Analyses



## (0) Demographic and clinical Info

### (0.1) summarise
```{r,echo=FALSE,warning=FALSE}
# Raw data
par(mfrow=c(2,4))
hist(df$age_m,las=1,main="Age_m",xlab=NULL,ylab="Count")
hist(df$age_d,las=1,main="Age_d",xlab=NULL,ylab="Count")
hist(df$drinking_class,las=1,main="Drinking_class",xlab=NULL,ylab="Count")
barplot(table(df$sex),space=0.3,col="white",las=1,main="Sex",xlab=NULL,ylab="Count")
barplot(table(df$race_label),horiz=F,col="white",las=2,main="Race",xlab=NULL,ylab="Count",na.action=T)
hist(df$ses,las=1,main="Ses",xlab=NULL,ylab="Count")
hist(df$fh_alc_density,las=1,main="fh_alc_density",xlab=NULL,ylab="Count")
hist(df$life_trauma_RP,las=1,main="Life_trauma",xlab=NULL,ylab="Count")


# Log10() transformed data (suggested by Raj)
par(mfrow=c(2,3))
hist(df$ses,las=1,main="Ses",xlab=NULL,ylab="Count")
hist(df$fh_alc_density,las=1,main="fh_alc_density",xlab=NULL,ylab="Count")
hist(df$life_trauma_RP,las=1,main="Life_trauma",xlab=NULL,ylab="Count")

hist(log10(df$ses),las=1,main="log10(Ses)",xlab=NULL,ylab="Count")
hist(log10(df$fh_alc_density),las=1,main="log10(df$fh_alc_density)",xlab=NULL,ylab="Count")
hist(log10(df$life_trauma_RP),las=1,main="log10(Life_trauma)",xlab=NULL,ylab="Count")

# Transformed data (suggested by Mike)
# race info by Mike
df_race <- read.csv("NCANDA_raceEdata.csv",header=T) %>% mutate(subject=ID,baseline=race3,followup_1y=race3,followup_2y=race3,followup_3y=race3) %>% select(c("subject","baseline","followup_1y","followup_2y","followup_3y")) %>% pivot_longer(-subject, names_to = "visit", values_to = "race_label")
df <- merge(df,df_race,by=c("subject","visit"),all.x=T) %>% rename(race_label=race_label.y)

df <- df %>% mutate(drinking_class = replace(drinking_class,drinking_class==3,2))
df <- df %>% mutate(ses = replace(ses,ses<=12,0)) %>% mutate(ses = replace(ses,ses>12,1))
df <- df %>% mutate(life_trauma_RP = replace(life_trauma_RP,life_trauma_RP==5,4))
df <- df %>% mutate(race_label = str_replace_all(race_label,c("Caucasian/White"="Cauc","African-American/Black"="AAmer","other"="Other")))

par(mfrow=c(2,4))
hist(df$age_m,las=1,main="Age_m",xlab=NULL,ylab="Count")
hist(df$age_d,las=1,main="Age_d",xlab=NULL,ylab="Count")
hist(df$drinking_class,las=1,main="Drinking_class",xlab=NULL,ylab="Count")
barplot(table(df$sex),space=0.3,col="white",las=1,main="Sex",xlab=NULL,ylab="Count")
barplot(table(df$race_label),horiz=F,col="white",las=2,main="Race",xlab=NULL,ylab="Count",na.action=T)
hist(df$ses,las=1,main="Ses",xlab=NULL,ylab="Count")
hist(df$fh_alc_density,las=1,main="fh_alc_density",xlab=NULL,ylab="Count")
hist(df$life_trauma_RP,las=1,main="Life_trauma",xlab=NULL,ylab="Count")

print("Number of subjects grouped by drinking_class")
df %>% filter(visit=="baseline") %>% select(drinking_class) %>% group_by(drinking_class) %>% summarise(n=n())
print("Number of subjects grouped by ses")
df %>% filter(visit=="baseline") %>% select(ses) %>% group_by(ses) %>% summarise(n=n())

```


#Table 1.
```{r,echo=FALSE}
# number of subjects
dfx <- df %>% select(c('subject','visit','drinking_class')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% filter(baseline==0) # 657 subjects with baseline drinking_class=0
dfx %>% summarise(n=n())
dfx %>% filter(followup_1y>=0) %>% summarise(n=n()) # 576 subjects in followup_1y
dfx %>% filter(followup_2y>=0) %>% summarise(n=n()) # 536 subjects in followup_2y
dfx %>% filter(followup_3y>=0) %>% summarise(n=n()) # 484 subjects in followup_3y
 
# number of subjects per drinking_class per visit
dfx <- df %>% select(c('subject','visit','drinking_class')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% filter(baseline==0) # 657 subjects with baseline drinking_class=0
dfx %>% select('baseline') %>% table()
dfx %>% select('followup_1y') %>% table()
dfx %>% select('followup_2y') %>% table()
dfx %>% select('followup_3y') %>% table()

# number of subjects per sex
dfx <- df %>% select(c('subject','visit','drinking_class','sex')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,sex)) %>% filter(drinking_class_baseline==0) # 657 subjects with baseline drinking_class=0
dfx %>% group_by(sex_baseline) %>% summarise(n=n())
dfx %>% filter(drinking_class_followup_1y>=0) %>% group_by(sex_followup_1y) %>% summarise(n=n()) 
dfx %>% filter(drinking_class_followup_2y>=0) %>% group_by(sex_followup_2y) %>% summarise(n=n()) 
dfx %>% filter(drinking_class_followup_3y>=0) %>% group_by(sex_followup_3y) %>% summarise(n=n()) 
 
# number of subjects per race_label
dfx <- df %>% select(c('subject','visit','drinking_class','race_label')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,race_label)) %>% filter(drinking_class_baseline==0) # 657 subjects with baseline drinking_class=0
dfx %>% group_by(race_label_baseline) %>% summarise(n=n())
dfx %>% filter(drinking_class_followup_1y>=0) %>% group_by(race_label_followup_1y) %>% summarise(n=n()) 
dfx %>% filter(drinking_class_followup_2y>=0) %>% group_by(race_label_followup_2y) %>% summarise(n=n()) 
dfx %>% filter(drinking_class_followup_3y>=0) %>% group_by(race_label_followup_3y) %>% summarise(n=n()) 
#dfx <- df %>% select(c('subject','visit','drinking_class','race_label')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,race_label)) %>% filter(drinking_class_baseline==0) 
#SDL_calc <- function(col){col %>% as.factor() %>% table(useNA="no")}
#dfx %>% select("race_label_baseline":"race_label_followup_3y") %>% sapply(SDL_calc) # works for number of subjects
#dfx$race_label_followup_3y %>% base::table(exclude=c(NA,NaN),useNA="no")
 
# age per visit
dfx <- df %>% select(c('subject','visit','drinking_class','mri_t1_age')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,mri_t1_age)) %>% filter(drinking_class_baseline==0)
dfx %>% select(mri_t1_age_baseline:mri_t1_age_followup_3y) %>% summarise_all(list(~mean(.,na.rm=TRUE),~sd(.,na.rm=TRUE)))
 
# SES per visit
dfx <- df %>% select(c('subject','visit','drinking_class','ses')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,ses)) %>% filter(drinking_class_baseline==0) # 657 subjects with baseline drinking_class=0
dfx %>% group_by(ses_baseline) %>% summarise(n=n())
dfx %>% filter(drinking_class_followup_1y>=0) %>% group_by(ses_followup_1y) %>% summarise(n=n()) 
dfx %>% filter(drinking_class_followup_2y>=0) %>% group_by(ses_followup_2y) %>% summarise(n=n()) 
dfx %>% filter(drinking_class_followup_3y>=0) %>% group_by(ses_followup_3y) %>% summarise(n=n()) 
 
# fh_alc_density per visit
dfx <- df %>% select(c('subject','visit','drinking_class','fh_alc_density')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,fh_alc_density)) %>% filter(drinking_class_baseline==0)
dfx %>% select(fh_alc_density_baseline:fh_alc_density_followup_3y) %>% summarise_all(list(~mean(.,na.rm=TRUE),~sd(.,na.rm=TRUE))) 
 
# life_trauma per visit
dfx <- df %>% select(c('subject','visit','drinking_class','life_trauma_RP')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,life_trauma_RP)) %>% filter(drinking_class_baseline==0)
dfx %>% select(life_trauma_RP_baseline:life_trauma_RP_followup_3y) %>% summarise_all(list(~mean(.,na.rm=TRUE),~sd(.,na.rm=TRUE))) 

```

# Supplementary Table 1.
```{r,echo=FALSE}
# number of subjects
dfx <- df %>% select(c('subject','visit','drinking_class','site')) %>% pivot_wider(names_from = visit, values_from = drinking_class) %>% filter(baseline==0) # 653 subjects with baseline drinking_class=0
dfx %>% group_by(site) %>% summarise(n=n())

# number of subjects per sex
dfx <- df %>% select(c('subject','visit','drinking_class','site','sex')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,sex)) %>% filter(drinking_class_baseline==0) # 653 subjects with baseline drinking_class=0
dfx %>% group_by(site,sex_baseline) %>% summarise(n=n())
 
# age at baseline
dfx <- df %>% select(c('subject','visit','drinking_class','mri_t1_age','site')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,mri_t1_age)) %>% filter(drinking_class_baseline==0)
dfx %>% select(c(site,mri_t1_age_baseline)) %>% group_by(site) %>%  summarise_all(list(~mean(.,na.rm=TRUE),~sd(.,na.rm=TRUE)))

# number of subjects per race_label at baseline
dfx <- df %>% select(c('subject','visit','drinking_class','race_label','site')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,race_label)) %>% filter(drinking_class_baseline==0) # 653 subjects with baseline drinking_class=0
dfx %>% group_by(site,race_label_baseline) %>% summarise(n=n())

 
# SES at baseline
dfx <- df %>% select(c('subject','visit','drinking_class','ses','site')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,ses)) %>% filter(drinking_class_baseline==0) # 653 subjects with baseline drinking_class=0
dfx %>% group_by(site,ses_baseline) %>% summarise(n=n())
 
 
# fh_alc_density at baseline
dfx <- df %>% select(c('subject','visit','drinking_class','fh_alc_density','site')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,fh_alc_density)) %>% filter(drinking_class_baseline==0)
dfx %>% select(c(site,fh_alc_density_baseline)) %>% group_by(site) %>% summarise_all(list(~mean(.,na.rm=TRUE),~sd(.,na.rm=TRUE))) 
 
# life_trauma at baseline
dfx <- df %>% select(c('subject','visit','drinking_class','life_trauma_RP','site')) %>% pivot_wider(names_from = visit, values_from = c(drinking_class,life_trauma_RP)) %>% filter(drinking_class_baseline==0)
dfx %>% select(c(site,life_trauma_RP_baseline)) %>% group_by(site) %>% summarise_all(list(~mean(.,na.rm=TRUE),~sd(.,na.rm=TRUE))) 

```





## (1) Main effect of drinking class, age_m and age_d

### (1.1) main effect of drinking_class
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_m + age_d + drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("drinking_class"),fxy=c("Drinking Class","Cortical Thickness"),ffact='drinking_class')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="drinking_class"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter 

```

### (1.2) main effect of age_d
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_m + age_d + drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_d"),fxy=c("age_d","Mean Value"),ffact='age_d')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_d"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter 

```


### (1.3) main effect of age_m
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_m + age_d + drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_m"),fxy=c("age_m","Mean Value"),ffact='age_m')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_m"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter 

```



## (2) Two-way interactions of drinking class and the other fixed variables

## (2.1) drinking class x age_d interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_m + age_d * drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_d","drinking_class"),fxy=c("Within-Person Age Change (Years)","Mean Value"),ffact='age_d:drinking_class')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_d:drinking_class"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```





## (2.2) drinking class x age_m interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + age_m * drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_m","drinking_class"),fxy=c("age_m","Mean Value"),ffact='age_m:drinking_class')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_m:drinking_class"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```


## (2.3) age_d x age_m interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d * age_m + drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_d","age_m"),fxy=c("age_d","Mean Value"),ffact='age_d:age_m')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_d:age_m"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```

## (2.4) drinking class x life_trauma interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + age_m + drinking_class * life_trauma_RP + sex + race_label + ses + fh_alc_density + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("drinking_class","life_trauma_RP"),fxy=c("Drinking Class","Mean Value"),ffact='drinking_class:life_trauma_RP')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="drinking_class:life_trauma_RP"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```

## (2.5) drinking class x sex interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + age_m + drinking_class * sex + life_trauma_RP + race_label + ses + fh_alc_density + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("drinking_class","sex"),fxy=c("Drinking Class","Mean Value"),ffact='drinking_class:sexM')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="drinking_class:sexM"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```


## (2.6) drinking class x rae_label interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + age_m + drinking_class * race_label + sex + life_trauma_RP + ses + fh_alc_density + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("drinking_class","race_label"),fxy=c("Drinking Class","Mean Value"),ffact='drinking_class:race_labelCauc')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="drinking_class:race_labelCauc"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```



## (2.7) drinking class x ses interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + age_m + drinking_class * ses + race_label + sex + life_trauma_RP + fh_alc_density + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("drinking_class","ses"),fxy=c("Drinking Class","Mean Value"),ffact='drinking_class:ses')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="drinking_class:ses"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```


## (2.8) drinking class x fh_alc_density interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + age_m + drinking_class * fh_alc_density + ses + race_label + sex + life_trauma_RP + (1|subject))" # defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("drinking_class","fh_alc_density"),fxy=c("Drinking Class","Mean Value"),ffact='drinking_class:fh_alc_density')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="drinking_class:fh_alc_density"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```




## (3) 3-Way Interactions

## (3.1) age_d x age_m x drinking_class interaction (with 2-way interactions)
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d * age_m * drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # !!! defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_d","age_m","drinking_class"),fxy=c("Within-Person Age Change (Years)","Mean Value"),ffact='age_d:age_m:drinking_class')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_d:age_m:drinking_class"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter
```

## (3.2) age_d x age_m x drinking_class interaction (without 2-way interactions)
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d:age_m:drinking_class + age_d + age_m + drinking_class + sex + race_label + ses + fh_alc_density + life_trauma_RP + (1|subject))" # !!! defined by the user

# all functions
df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_d","age_m","drinking_class"),fxy=c("age_d","Mean Value"),ffact='age_d:age_m:drinking_class')

# coefficients
coefs <- df %>% select(c1:c7) %>% sapply(SDL_lm, df=df, Otype="coef", ftxt=ftxt) #coefs is a list of matrix

# coefficients of interest
coef_inter <- coefs %>% do.call(rbind,.) %>% .[which(rownames(.)=="age_d:age_m:drinking_class"),] %>% as.data.frame() %>% mutate(p_fdr=p.adjust(.[,"Pr(>|t|)"],method="fdr"))
row.names(coef_inter) <- sprintf('c%02d',1:7)

coef_inter

```












# (2) Effects of life_trauma

## (2.1) life_trauma main effect
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_m + age_d + drinking_class + life_trauma_RP + sex + race_label + ses + fh_alc_density + (1|subject))" # !!! defined by the user

df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("life_trauma_RP"),fxy=c("Life_trauma","Mean Value"),ffact='life_trauma_RP')

```



## (2.2) age_m x life_trauma interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_d + drinking_class + age_m * life_trauma_RP + sex + race_label + ses + fh_alc_density + (1|subject))" # !!! defined by the user

df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_m","life_trauma_RP"),fxy=c("Age_m","Mean Value"),ffact='age_m:life_trauma_RP')

```




## (2.3) age_d x life_trauma interaction
```{r,echo=FALSE}
# define model formula
ftxt <- "lmer(formula = y ~ age_m + drinking_class + age_d * life_trauma_RP + sex + race_label + ses + fh_alc_density + (1|subject))" # !!! defined by the user

df %>% select(c1:c7) %>% SDL_all(df=df,fcorr='fdr',ftxt=ftxt,fterms=c("age_d","life_trauma_RP"),fxy=c("Age_d","Mean Value"),ffact='age_d:life_trauma_RP')

```
